\chapter{The Gap: WHy Does It Exists}\label{chapter:gap}

Before diving into the details of bridging the gap between local code development
and cloud execution in \ac{MLOps}, it is essential to understand how the gap exists in the first place.
In this chapter, I will explore a brief explanation of why there is a gap between the two environments.

\section{Why the Gap Exists}
There is a notable distinction between running code locally on a developer's machine and deploying it to a cloud server for execution. This disparity can pose certain challenges in terms of infrastructure, environment, configuration and more, which can make it challenging to ensure seamless functionality across both environments.
\newline 
I would like to take this opportunity to examine a few of the reasons that have been presented. It is important to consider these reasons in a fair and impartial way:
\begin{itemize}
    \item Local development environments often have different configurations, dependencies, and resources than cloud environments. This can lead to inconsistencies and incompatibilities when deploying code to the cloud.
    \item Cloud environments often have different security, compliance, and scalability requirements than local environments. This can lead to additional complexity and overhead when deploying code to the cloud.
    \item Cloud environments often have different monitoring, logging, and debugging tools than local environments. This can make it difficult to diagnose and resolve issues that arise in the cloud.
\end{itemize}
In the coming chapters, I will discuss different approaches to overcoming these difficulties and how they
can be implemented in practice and put a spurt in the development and debugging
process on the cloud.


